Planning to Save Energy: How Information Format Affects Accuracy

### 2025-03-31 manuscript/manuscript.qmd
major edits for natcom

---
title: "Planning to Save Energy: How Information Format Affects Accuracy"
aliases:
  - /manuscript.html
format-links:   
    - text: HTML
      href: manuscript.html
      icon: file-minus
    - text: HTML (with code)
      href: manuscript-code.html
      icon: file-plus
    - format: apaquarto-pdf
      text: PDF version
      icon: file-pdf
    - text: docx
      href: rs_manuscript_apa.docx
    - gfm
date: 2025-03-16
abstract: "Effective communication of energy consumption information is crucial for promoting residential energy conservation. This study investigates how different numerical representations of energy reduction goals influence consumers' ability to create accurate conservation plans. Across two experiments, we examined the impact of presenting energy information in kilowatt-hours (kWh), percentages, or U.S. dollars (USD) on planning accuracy. Participants completed a simulated household planning task in which they allocated energy usage across multiple appliances, with the goal presented in either kilowatt-hours (kWh), percentages, or monetary costs. Results across both experiments showed that presenting reduction goals in absolute units (kWh) led to significantly greater accuracy compared to percentage-based or monetary formats. Furthermore, we found that higher energy literacy was associated with more accurate planning. These findings demonstrate that absolute units (kWh) are more effective for communicating energy-saving goals, and highlight the potential value of educational interventions to improve consumer energy literacy."
---

@@

Receiving feedback about one’s energy use is important in identifying potential actions to reduce energy consumption. However, it is crucial to understand how consumers use the information provided with their energy bill and whether they are able to translate an energy-saving goal into an action plan. @canfieldPerceptionsElectricityuseCommunications2017 specifically studied this problem by running an experiment in which participants were shown hypothetical electricity bills with information related to a household’s historical electricity use, electricity use in relation to their neighbors, and historical electricity use by appliances. Participants were randomly assigned to one of three formats of information representation (i.e., tables, bar graphs, and icon graphs) and were asked questions regarding their energy literacy. @canfieldPerceptionsElectricityuseCommunications2017 showed that tables were the easiest format to understand for consumers when evaluating every type of information related to energy use. Across types of information, historical electricity use elicited the highest intentions and preferences for energy savings regardless of format. Additionally, participants with high energy literacy had a better understanding of energy-related information across all types of information representations. By disentangling the effects of content, format, and individual differences in energy literacy on understanding, preferences, and intentions, @canfieldPerceptionsElectricityuseCommunications2017 demonstrated that easy-to-implement communication strategies in energy bills can lead to energy-saving behaviors.

@@

Similarly, @tonkeProvidingProceduralKnowledge2024 rreported that sending households brief but precise text messages outlining how to reduce water use (e.g., limiting irrigation times, adjusting washing machine settings) yielded meaningful decreases in consumption, underscoring the importance of procedural knowledge—namely, knowing how to operationalize a goal rather than simply why it is desirable. In the context of energy conservation, this implies that interventions should not only highlight potential reductions (such as a 15% target) but also guide residents in allocating those reductions across specific appliances or behaviors [@attariPublicPerceptionsEnergy2010]. Additionally, meta-analytic findings suggest that people respond more robustly to household-level feedback that situates their usage within a personalized framework, thereby reducing the cognitive burden of determining next steps [@nematiImpactInformationbasedInterventions2020].

@@

Building on the study of Canfield et al. (2017), we aimed to test xxx.

@@

We implemented our task and surveys on Qualtrics, and recruited participants through Amazon Mechanical Turk. In Experiment 1, 235 participants were initially recruited. Data from six participants were excluded due to deviant performance on the task, resulting in a final sample of 229 participants (146 males, 92 females, 1 not specified). The average age of participants was 34.3 years (SD = 10.2). Most participants (76%) reported using a calculator to complete the task.

@@

The study employed a mixed design with reference class (kWh, percentage, USD) as a between-subjects factor and state/family scenario as a within-subjects factor. Each participant completed energy reduction planning tasks for two different states (i.e., family scenarios), with state order counterbalanced across participants. The family scenarios featured four households in different climate regions: Texas (Smith family) and California (Adams family) representing warm climates, and Colorado (Wells family) and Massachusetts (Davis family) representing cold climates. Average utility use data for each state was obtained from the  CITE SOURCE FOR STATE AVGS? Participants also completed an 8-item questionnaire assessing their knowledge of energy consumption and conversion [@dewatersEnergyLiteracySecondary2011], and a question indicating whether they used a calculator for the task.

@@

![Example energy planning task trial](/assets/images/wells10e_pub.png){#fig-task width="80%" apa-note="Participants saw a table with a family's previous year electricity usage (here for the Wells family in Colorado) and were asked to allocate energy usage to meet a 15% reduction goal. The format of the reduction goal was manipulated to be either a percentage (15% given as goal reduction), kilowatt hours (5965 kWh given), or USD (\$656). Participants in the USD condition were provided with the conversion rate between kwH and USD."}

## Measures

@@

All preprocessing and analyses were carried out in R [@rcoreteamLanguageEnvironmentStatistical2020] and the Tidyverse package [@wickhamWelcomeTidyverse2019]. Mixed Bayesian regressions were fit using the brms package [@burknerBrmsPackageBayesian2017], with participants and family scenario (states) set as random effects.

@@

s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,edu,pct_goal,calc) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = abs(round((orig_kWh-total_kWh)/orig_kWh,3)), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(duration), .groups = 'drop') |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007), 
                close_match = abs_error <= 0.02) |>
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.05 ~ "0.01-5% error",
                    TRUE ~ "Over 5% error" 
                ), 
                levels = c("Exact match", "0.01-5% error", "Over 5% error"),
                ordered = TRUE
            )
        ) |> relocate(accuracy_level, .after= "pct_change")

s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(
    mg=sum(matched_goal),
    mgc=sum(close_match),
    n=n(), 
    pct=mg/n,
    pct_close=mgc/n,
    mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),
    mean_log_abs_error=mean(log_abs_error)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.05 ~ "0.01-5% error",
                TRUE ~ "Over 5% error" 
            ), 
            levels = c("Exact match", ".01-5% error", "Over 5% error"),
            ordered = TRUE
        ))

@@

# overall pct of subjects who matched their goal
# s1_agg4 |> group_by('Reference Class' = refClass) |>
#     summarise(
#     'Avg. % Change' = mean(mean_pct_change),
#     '% meeting goal (exact)' = mean(pct),
#     '% meeting goal (close match)' = mean(pct_close),
#     'Abs. Deviation' = median(mean_abs_error),
#     'Log Abs. Deviation' = (median(mean_log_abs_error)),
#     # sd = sd(pct),
#     # n = n(),
#     #se=sd(pct)/sqrt(n)
# ) |>   mutate(across(where(is.numeric), \(x) round(x, 3))) %>% 
#   kable(escape=FALSE,booktabs=TRUE,align=c("l"))

#pander::pandoc.table(caption="Study 1: Proportion of participants who matched their goal overall")

summary_table <- s1_agg4 |> 
  group_by(`Reference Class` = refClass) |> 
  summarise(
    N = n(),
    `Avg. % Change` = sprintf("%.2f (%.2f)", mean(mean_pct_change), sd(mean_pct_change)),
    `% meeting goal (exact)` = sprintf("%.2f (%.2f)", mean(pct), sd(pct)),
    `% meeting goal (close match)` = sprintf("%.2f (%.2f)", mean(pct_close), sd(pct_close)),
    `Abs. Deviation` = sprintf("%.2f (%.2f)", mean(mean_abs_error), sd(mean_abs_error)),
    `Log Abs. Deviation` = sprintf("%.2f (%.2f)", mean(mean_log_abs_error), sd(mean_log_abs_error))
  ) |> 
  kable(escape = FALSE, booktabs = TRUE, align = c("l"))

@@

::: {#tbl-s1-agg apa-note="Summary statistics by reference class. Values represent mean (SD) across participants for each planning accuracy metric. Values in parentheses represent standard deviations. 'Close match' indicates trials where the proposed reduction was within ±2% of the target goal. Abs. Deviation and Log Abs. Deviation columns report medians."}

|Reference Class |N  |Avg. % Change |% meeting goal (exact) |% meeting goal (close match) |Abs. Deviation |Log Abs. Deviation |
|:---------------|:--|:-------------|:----------------------|:----------------------------|:--------------|:------------------|
|kWh             |76 |0.22 (0.14)   |0.38 (0.45)            |0.54 (0.45)                  |0.09 (0.14)    |-3.47 (1.43)       |
|Percentage      |67 |0.21 (0.11)   |0.22 (0.37)            |0.40 (0.41)                  |0.09 (0.10)    |-3.18 (1.20)       |
|USD             |86 |0.23 (0.13)   |0.10 (0.28)            |0.22 (0.33)                  |0.12 (0.11)    |-2.61 (1.06)       |

@@

@tbl-s1-agg presents a summary of descriptive statistics regarding planning accuracy across the three experimental reference class conditions (kWh, Percentage, USD). Observation of the central tendencies suggests a potential advantage for the kWh condition. The kWh group exhibited the highest mean proportion of trials exactly matching the target goal (M = 0.38, SD = 0.45), considerably higher than that observed in the Percentage (M = 0.22, SD = 0.37) and USD (M = 0.10, SD = 0.28) conditions. A concordant pattern emerged for deviations, where the kWh condition displayed the lowest median absolute deviation (Mdn = 0.03) compared to the Percentage (Mdn = 0.06) and USD (Mdn = 0.10) conditions, hinting at greater precision when goals were presented in absolute energy units.

@@

```{r}
#| label: tbl-s1-prop
#| tbl-cap: "Experiment 1 1: Categorization of Participants according to Accuracy Levels"
#| apa-note: "The table shows the percentage of participants who fell into each accuracy level for each reference class condition (percentages of kWh, $, and USD columns reflect within condition percentages). The combined group column reflects the percentage of participants in each accuracy level when aggregating across across all reference class conditions."

@@

observed_props_s1 |> 
  mutate(n_total=sum(n)/4) |> 
  group_by(accuracy_level) |>
  mutate(ns=sum(n)/4) |> 
  mutate(Total = paste0(round(ns/n_total*100,1), "%")) |>
  select('Reference Class'=refClass, 'Accuracy Level'=accuracy_level, '% in Group'=pct_grp, "Combined Groups %" =Total) |>
  pivot_wider(
    names_from = 'Reference Class',
    values_from = c('% in Group')
  ) |> relocate("Combined Groups %" , .after=last_col()) |> 
  kable(escape=FALSE,booktabs=TRUE,align=c("l"))

@@

We next categorized responses into three accuracy levels (exact match [0% error], minor deviations [0.01–5%], and large deviations [>5%]) for our primary statistical modeling. Using Bayesian ordinal regression, we modeled the ordered accuracy outcome as a function of the reference class condition, while controlling for random variation across participants and family scenarios:

$$
\text{Accuracy Level} \sim \text{Reference Class} + \text{Calculator} + (1|\text{id}) + (1|\text{Family Scenario})
$$

This approach allowed us to estimate thresholds (intercepts) and regression coefficients that capture how different reference classes affect the likelihood of achieving higher accuracy categories. For each comparison, we provide posterior odds ratios (OR) and their 95% CIs. This approach allows the estimation of threshold parameters and regression coefficients that characterize how changes in predictor variables (such as the reference class: kWh, percentage, or USD) relate to probabilities of being in each accuracy category. Specifically, we used a cumulative logit link function to model the ordered accuracy outcome, and we specified weakly informative priors for the regression coefficients (normal distributions with mean 0 and standard deviation of 1) and for the cutpoints (normal distributions with a mean of zero and a standard deviation of 4.0). The approach can be used to estimate threshold parameters and regression coefficients that characterize how changes in predictor variables (such as the reference class: kWh, percentage, or USD) relate to probabilities of being in each accuracy category.

@@

::: {#tbl-s1-reg apa-note="Ordinal regression results. Positive coefficients for the reference class predictors indicate that those conditions are associated with higher error categories relative to the kWh baseline."}

|Parameter          |Estimate |CI_Lower |CI_Upper |pd   |
|:------------------|:--------|:--------|:--------|:----|
|Intercept[1]       |-4.21    |-5.90    |-2.58    |1.00 |
|Intercept[2]       |-0.89    |-2.49    |0.71     |0.87 |
|refClassPercentage |1.44     |0.07     |2.88     |0.98 |
|refClassUSD        |3.13     |1.81     |4.50     |1.00 |
|calcUsedCalculator |-3.30    |-4.80    |-1.92    |1.00 |

@@

As shown in @tbl-s1-reg, the reference class coefficients are positive for both the Percentage (Estimate = 1.44, 95% CI: 0.07 to 2.88, pd = 0.98) and USD (Estimate = 3.13, 95% CI: 1.81 to 4.50, pd = 1.00) conditions, relative to the kWh baseline. This indicates that, compared to the kWh condition, participants in both the Percentage and USD conditions were more likely to produce plans that fell into higher error categories. Moreover, the odds ratios (see Table 3) suggest that the USD condition led to a notably higher likelihood of large errors compared to the kWh baseline (OR = 22.9), while the Percentage condition also demonstrated increased odds (OR = 4.2) but was somewhat less detrimental to accuracy than USD. These results align with our descriptive findings and further clarify that framing the target reductions in absolute kWh units may facilitate significantly more accurate planning. Posterior predictive checks showed that the ordinal model provided a reasonable fit to the observed data (see @fig-s1-ppd).

```{r}
#| label: fig-s1-ppd
#| fig-cap: "Experiment 1: Posterior Predictive Check"
#| apa-note: "Posterior predictive check for frequency of trials at each accuracy level, faceted by reference class (kWh, Percentage, USD). Bars show observed, dots show model predicted proportions. Better accuracy is indicated by higher proportions in the 'Exact Match' and '0.01-5% error' categories."

@@

To further investigate individual factors that may influence planning accuracy, we examined the relationship between participants' energy literacy scores and their performance on the task. Energy literacy was assessed using an 8-item questionnaire adapted from [@dewatersEnergyLiteracySecondary2011], which covers topics such as energy units, appliance energy consumption, and sources of electricity. A Bayesian linear regression model was fit with log-transformed absolute error as the outcome variable and energy literacy score as the predictor, controlling for random effects of participant and state: log_abs_error ~ els + (1|id) + (1|state). Results indicated a significant negative relationship between energy literacy and log absolute error (Estimate = -2.35, 95% CI: -2.88 to -1.81), suggesting that participants with higher energy literacy scores tended to have smaller deviations from the target reduction goal, and thus more accurate plans overall (@fig-s1-els).

@@

Experiment 1 examined how different numerical representations of energy reduction goals affected participants’ planning accuracy. In line with our hypothesis that absolute units would yield better accuracy, the kWh condition supported significantly more precise energy reduction plans than did either the Percentage or USD conditions. Although the Percentage format was detrimental to accuracy relative to kWh, it was the USD condition that consistently produced the poorest outcomes, suggesting that monetary terms, while intuitive in everyday contexts, may not serve as effective reference classes for planning appliance-specific reductions in energy use.

@@

@@

In Experiment 2, We recruited 206 participants from Amazon Mechanical Turk, but data from 10 participants were corrupted due to experimenter error, and six excluded due to deviant behavior, leaving a final sample of 190 participants (102 male; 88 female. Average age = 35.5, SD=9.5)

Experiment 2 employed a mixed experimental design with reference class (USD, Percentage, kWh) as a between-subjects factor,  and two within-subjects factors: task goal (10% vs. 15% reduction) and the presentation of last year's usage data (exact vs. rounded numbers). The order of presentation of the goal, rounding, and state conditions was counterbalanced across participants.  As in Experiment 1, each participant completed the energy reduction planning task for two different family-state scenarios. In the “rounded” condition, both the family’s previous year usage and the state averages were rounded to the nearest whole number.

@@

@@

::: {#tbl-s2-agg apa-note="The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal. Percentage meeting goal refers to participants who exactly met the target reduction goal (i.e., the proposed plan would reduce energy use by exactly 15%)"}

>>>>

---
title: "Planning to Save Energy: How Information Format Affects Accuracy"
aliases:
  - /manuscript.html
format-links:   
    - text: HTML
      href: manuscript.html
      icon: file-minus
    - text: HTML (with code)
      href: manuscript-code.html
      icon: file-plus
    - format: apaquarto-pdf
      text: PDF version
      icon: file-pdf
    - text: docx
      href: rs_manuscript_apa.docx
    - gfm
date: 2025-03-31
keywords: 
  - energy literacy
  - information format
  - planning accuracy
  - decision making
abstract: "Effective communication of energy consumption information is crucial for promoting residential energy conservation. This study investigates how different numerical representations of energy reduction goals influence consumers' ability to create accurate conservation plans. Across two experiments, we examined the impact of presenting energy information in kilowatt-hours (kWh), percentages, or U.S. dollars (USD) on planning accuracy. Participants completed a simulated household planning task in which they allocated energy usage across multiple appliances, with the goal presented in either kilowatt-hours (kWh), percentages, or monetary costs. Results across both experiments showed that presenting reduction goals in absolute units (kWh) led to significantly greater accuracy compared to percentage-based or monetary formats. Furthermore, we found that higher energy literacy was associated with more accurate planning. These findings demonstrate that absolute units (kWh) are more effective for communicating energy-saving goals, and highlight the potential value of educational interventions to improve consumer energy literacy."
---

@@

Receiving feedback about one’s energy use is important in identifying potential actions to reduce energy consumption. However, it is crucial to understand how consumers use the information provided with their energy bill and whether they are able to translate an energy-saving goal into an action plan. @canfieldPerceptionsElectricityuseCommunications2017 specifically studied this problem by running an experiment in which participants were shown hypothetical electricity bills with information related to a household’s historical electricity use, electricity use in relation to their neighbors, and historical electricity use by appliances. Participants were randomly assigned to one of three formats of information representation (i.e., tables, bar graphs, and icon graphs) and were asked questions regarding their energy literacy. @canfieldPerceptionsElectricityuseCommunications2017 showed that tables were the easiest format to understand for consumers when evaluating every type of information related to energy use. Across types of information, historical electricity use elicited the highest intentions and preferences for energy savings regardless of format. Additionally, participants with high energy literacy had a better understanding of energy-related information across all types of information representations. Studies such as @dewatersEnergyLiteracySecondary2011 have highlighted that while individuals may express concern for energy issues, their actual knowledge and behaviors may not align, underscoring the importance of fostering a comprehensive understanding of energy concepts to facilitate effective conservation actions. By disentangling the effects of content, format, and individual differences in energy literacy on understanding, preferences, and intentions, @canfieldPerceptionsElectricityuseCommunications2017 demonstrated that easy-to-implement communication strategies in energy bills can lead to energy-saving behaviors.

@@

Similarly, @tonkeProvidingProceduralKnowledge2024 reported that sending households brief but precise text messages outlining how to reduce water use (e.g., limiting irrigation times, adjusting washing machine settings) yielded meaningful decreases in consumption, underscoring the importance of procedural knowledge—namely, knowing how to operationalize a goal rather than simply why it is desirable. In the context of energy conservation, this implies that interventions should not only highlight potential reductions (such as a 15% target) but also guide residents in allocating those reductions across specific appliances or behaviors [@attariPublicPerceptionsEnergy2010]. Additionally, meta-analytic findings suggest that people respond more robustly to household-level feedback that situates their usage within a personalized framework, thereby reducing the cognitive burden of determining next steps [@nematiImpactInformationbasedInterventions2020].

@@

@@

We implemented our task and surveys on Qualtrics, and recruited participants through Amazon Mechanical Turk. In Experiment 1, 235 participants were initially recruited. Data from six participants were excluded due to unusually large error magnitudes (error magnitudes exceeding 2.5 standard deviations from the group mean), resulting in a final sample of 229 participants (136 males, 92 females, 1 participant not specified). The average age of participants was 34.3 years (SD = 10.2). Most participants (76%) reported using a calculator to complete the task.

@@

The study employed a mixed design with reference class (kWh, percentage, USD) as a between-subjects factor and state/family scenario as a within-subjects factor. Each participant completed energy reduction planning tasks for two different states (i.e., family scenarios), with state order counterbalanced across participants. The family scenarios featured four households in different climate regions: Texas (Smith family) and California (Adams family) representing warm climates, and Colorado (Wells family) and Massachusetts (Davis family) representing cold climates. Participants also completed an 8-item questionnaire assessing their knowledge of energy consumption and conversion [@dewatersEnergyLiteracySecondary2011], and a question indicating whether they used a calculator for the task.

@@

![Example energy planning task trial](/assets/images/wells10e_pub.png){#fig-task width="80%" apa-note="Participants saw a table with a family's previous year electricity usage (here for the Wells family in Colorado) and were asked to allocate energy usage to meet a 15% reduction goal. The format of the reduction goal was manipulated to be either a percentage (15% reduction), kilowatt hours (5965 kWh), or USD (\$656). Participants in the USD condition were provided with the conversion rate between kWh and USD."}

@@

All preprocessing and analyses were carried out in `R` version 4.4.3 [@rcoreteamLanguageEnvironmentStatistical2020] and the Tidyverse package [@wickhamWelcomeTidyverse2019]. Mixed Bayesian regressions were fit using the `brms` package [@burknerBrmsPackageBayesian2017], with participants and family scenario set as random effects.

@@

```{r}
s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,edu,pct_goal,calc) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = abs(round((orig_kWh-total_kWh)/orig_kWh,3)), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(duration), .groups = 'drop') |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007), 
                close_match = abs_error <= 0.02) |>
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.05 ~ "0.01-5% error",
                    TRUE ~ "Over 5% error" 
                ), 
                levels = c("Exact match", "0.01-5% error", "Over 5% error"),
                ordered = TRUE
            )
        ) |> relocate(accuracy_level, .after= "pct_change")

s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(
    mg=sum(matched_goal),
    mgc=sum(close_match),
    n=n(), 
    pct=mg/n,
    pct_close=mgc/n,
    mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),
    mean_log_abs_error=mean(log_abs_error)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.05 ~ "0.01-5% error",
                TRUE ~ "Over 5% error" 
            ), 
            levels = c("Exact match", ".01-5% error", "Over 5% error"),
            ordered = TRUE
        ))
```

@@

summary_table <- s1_agg4 |>
    group_by(`Reference Class` = refClass) |>
    summarise(
        N = n(),
        # Keep Mean (SD) for average change proportion
        #`Avg. % Change` = sprintf("%.2f (%.2f)", mean(mean_pct_change), sd(mean_pct_change)),
        # Calculate Mean (SD) for percentages, multiply by 100, format as whole number
        `Exact Match Rate(%)` = sprintf("%.0f (%.0f)", mean(pct) * 100, sd(pct) * 100),
        `Close Match Rate(%)` = sprintf("%.0f (%.0f)", mean(pct_close) * 100, sd(pct_close) * 100),
        # Calculate Median [IQR] for deviations, adjust decimals as needed
        `Abs. Deviation` = sprintf("%.3f (%.3f)", median(mean_abs_error), IQR(mean_abs_error)),
        `Log Abs. Deviation` = sprintf("%.2f (%.2f)", median(mean_log_abs_error), IQR(mean_log_abs_error))
    ) |>
    # Apply kable formatting (optional, depends on your workflow)
    knitr::kable(escape = FALSE, booktabs = TRUE, align = c("l"))

@@

::: {#tbl-s1-agg apa-note="Summary statistics by reference class condition, based on participant-level averages. Exact Match Rate and Close Match Rate represent the mean percentage of trials (per participant) where the proposed reduction exactly matched or closely matched (±2% tolerance) the target goal (standard deviations in parentheses), respectively. Abs. Deviation and Log Abs. Deviation reflect the median (IQR) of each participant’s absolute deviation from the target goal."}

|Reference Class |N  |Exact Match Rate(%) |Close Match Rate(%) |Abs. Deviation |Log Abs. Deviation |
|:---------------|:--|:-------------------|:-------------------|:--------------|:------------------|
|kWh             |76 |38 (45)             |54 (45)             |0.029 (0.117)  |-3.69 (2.79)       |
|Percentage      |67 |22 (37)             |40 (41)             |0.056 (0.104)  |-3.10 (1.87)       |
|USD             |86 |10 (28)             |22 (33)             |0.095 (0.094)  |-2.40 (1.09)       |

@@

The summary statistics in @tbl-s1-agg suggest descriptive differences in planning accuracy across reference class conditions, with the kWh condition yielding the highest rates of exact matches, and lowest absolute deviations from the target reduction goal. However, use of these performance metrics to compare conditions is complicated by the non-normality of their distributions. Specifically, the substantial standard deviations associated with the match rates in @tbl-s1-agg, coupled with the visual evidence presented in @fig-s1-log-dist, underscore the markedly skewed, and potentially bi-modal distribution of the underlying planning errors.

In light of these distributional characteristics, we elected to bin planning performance into ordered accuracy levels (as shown in @tbl-s1-prop). Consequently, we binned individual trial outcomes into three distinct, ordered categories: "Exact match" (0% absolute error), "0.01-5% error" (minor deviations), and "Over 5% error" (larger deviations).  This binned `accuracy_level` variable subsequently served as the dependent measure in our primary inferential analyses examining the impact of reference class on planning accuracy

@@

```{r}
#| label: tbl-s1-prop
#| tbl-cap: "Experiment 1: Categorization of Participants according to Accuracy Levels"
#| apa-note: "The table shows the percentage of participants who fell into each accuracy level for each reference class condition (percentages of kWh, $, and USD columns reflect within condition percentages). "

# The combined group column reflects the percentage of participants in each accuracy level when aggregating across across all reference class conditions.

@@

observed_props_s1 |> 
  mutate(n_total=sum(n)/4) |> 
  group_by(accuracy_level) |>
  mutate(ns=sum(n)/4) |> 
  mutate(Total = paste0(round(ns/n_total*100,1), "%")) |>
  select('Reference Class'=refClass, 'Accuracy Level'=accuracy_level, '% in Group'=pct_grp) |> #, "Combined Groups %" =Total) |>
  pivot_wider(
    names_from = 'Reference Class',
    values_from = c('% in Group')
  ) |> #relocate("Combined Groups %" , .after=last_col()) |> 
  kable(escape=FALSE,booktabs=TRUE,align=c("l"))

@@

Group comparisons were conducted with Bayesian ordinal regression [@burknerOrdinalRegressionModels2019; @kruschkeDoingBayesianData2014]. We modeled the ordered accuracy outcome as a function of the reference class condition, while controlling for random variation across participants and family scenarios (@eq-e1).

$$
\text{Accuracy Level} \sim \text{Reference Class} + \text{Calculator} + (1|\text{id}) + (1|\text{Family Scenario})
$${#eq-e1}

This approach allows the estimation of threshold parameters and regression coefficients that characterize how changes in predictor variables (such as the reference class: kWh, percentage, or USD) relate to probabilities of being in each accuracy category. We used a cumulative logit link function to model the ordered accuracy outcome, and we specified weakly informative priors [@kruschkeDoingBayesianData2014] for the regression coefficients (normal distributions with mean 0 and standard deviation of 1) and for the cutpoints (normal distributions with a mean of zero and a standard deviation of 4.0). The approach can be used to estimate threshold parameters and regression coefficients that characterize how changes in predictor variables (such as the reference class: kWh, percentage, or USD) relate to probabilities of being in each level of accuracy (0% absolute error; 0.01-5% error;  Over 5% error). Bayesian regression provides a probability distribution over the plausible values of the model parameters, allowing for a probabilistic interpretation of the uncertainty associated with these estimates.The width of the credible intervals (Bayesian equivalent of confidence intervals) derived from the posterior distribution directly indicates the degree of uncertainty in these effects.

@@

::: {#tbl-s1-reg apa-note="Ordinal regression results. Positive coefficient estimates for the reference class predictors indicate that those conditions are associated with higher error categories relative to the kWh baseline."}

|Parameter          |Estimate |CI_Lower |CI_Upper |pd   |
|:------------------|:--------|:--------|:--------|:----|
|Intercept[1]       |-4.21    |-5.90    |-2.58    |1.00 |
|Intercept[2]       |-0.89    |-2.49    |0.71     |0.87 |
|RefClass-Percentage |1.44     |0.07     |2.88     |0.98 |
|RefClass-USD        |3.13     |1.81     |4.50     |1.00 |
|calc-UsedCalculator |-3.30    |-4.80    |-1.92    |1.00 |

@@

As shown in @tbl-s1-reg, the reference class coefficients are positive for both the Percentage (Estimate = 1.44, 95% CI: 0.07 to 2.88, pd = 0.98) and USD (Estimate = 3.13, 95% CI: 1.81 to 4.50, pd = 1.00) conditions, relative to the kWh baseline. This indicates that, compared to the kWh condition, participants in both the Percentage and USD conditions were more likely to produce plans that fell into higher error categories. Moreover, the odds ratios (see @tbl-s1-ord) suggest that the USD condition led to a notably higher likelihood of large errors compared to the kWh baseline (OR = 22.9), while the Percentage condition also demonstrated increased odds (OR = 4.2) but was somewhat less detrimental to accuracy than USD. These results align with our descriptive findings and further clarify that framing the target reductions in absolute kWh units may facilitate significantly more accurate planning. To evaluate the descriptive adequacy of our statistical model, we performed posterior predictive checks, a standard Bayesian technique wherein data simulated from the fitted ordinal model are compared against the observed data. As shown in @fig-s1-ppd, which compares the observed proportions of participant responses for each accuracy level (represented by bars) to the models predicted proportions (represented by points). The close correspondence between the observed bars and predicted points across different accuracy levels and reference class conditions suggests that the model adequately captures the patterns in the observed data.

```{r}
#| label: fig-s1-ppd
#| fig-cap: "Experiment 1: Posterior Predictive Check"
#| apa-note: "Posterior predictive check for the Bayesian ordinal regression model. Blue bars show the observed frequencies for each accuracy level (Exact match, 0.01-5% error, and Over 5% error), dots with error bars represent the model predictions. The plot is faceted by reference class condition (kWh, Percentage, USD)."

@@

To further investigate individual factors that may influence planning accuracy, we examined the relationship between participants' energy literacy scores and their performance on the task. A Bayesian linear regression model was fit with log-transformed absolute error as the outcome variable and energy literacy score as the predictor, controlling for random effects of participant and state. Results indicated a significant negative relationship between energy literacy and log absolute error (Estimate = -2.35, 95% CI: -2.88 to -1.81), suggesting that participants with higher energy literacy scores tended to have smaller deviations from the target reduction goal, and thus more accurate plans overall (@fig-s1-els).

@@

Experiment 1 examined how different numerical representations of energy reduction goals affected participants’ planning accuracy. In line with our hypothesis that absolute units would yield better accuracy, the kWh condition supported significantly more precise energy reduction plans than either the Percentage or USD conditions. Although the Percentage format was detrimental to accuracy relative to kWh, it was the USD condition that consistently produced the poorest outcomes, suggesting that monetary terms, while intuitive in everyday contexts, may not serve as effective reference classes for planning appliance-specific reductions in energy use.

@@

## Introduction

Building upon the preliminary evidence from Experiment 1, which indicated that representing energy reduction targets in absolute kilowatt-hour (kWh) units might yield superior planning accuracy relative to percentage or monetary formats, Experiment 2 was designed to further probe the robustness and nuances of this effect. A primary objective was to replicate the core finding regarding the influence of reference class on planning precision. Additionally, this second experiment sought to broaden the investigation by examining the potential impact of other task parameters; specifically, we manipulated the difficulty of the reduction goal (comparing a 10% versus a 15% target) and the granularity of the presented numerical usage data (using exact versus rounded figures).

@@

In Experiment 2, 206 participants were recruited from Amazon Mechanical Turk, but data from 10 participants were corrupted due to experimenter error, and six were excluded because their error magnitudes were statistical outliers (error magnitudes exceeding 2.5 standard deviations from the group mean), leaving a final sample of 190 participants (102 males, 88 females, with an average age = 35.5, SD=9.5)

Experiment 2 employed a mixed experimental design with reference class (USD, Percentage, kWh) as a between-subjects factor,  and two within-subjects factors: task goal (10% vs. 15% reduction) and the presentation of last year's usage data (exact vs. rounded numbers). The order of presentation of the goal, rounding, and family scenarios was counterbalanced across participants. As in Experiment 1, each participant completed the energy reduction planning task for two different family-state scenarios. In the “rounded” condition, both the family’s previous year usage and the state averages were rounded to the nearest whole number.

@@

summary_table <- s2_agg4 |>
    group_by(`Reference Class` = refClass) |>
    summarise(
        N = n(),
        # Keep Mean (SD) for average change proportion
        #`Avg. % Change` = sprintf("%.2f (%.2f)", mean(mean_pct_change), sd(mean_pct_change)),
        # Calculate Mean (SD) for percentages, multiply by 100, format as whole number
        `Exact Match Rate(%)` = sprintf("%.0f (%.0f)", mean(pct) * 100, sd(pct) * 100),
        `Close Match Rate(%)` = sprintf("%.0f (%.0f)", mean(pct_close) * 100, sd(pct_close) * 100),
        # Calculate Median [IQR] for deviations, adjust decimals as needed
        `Abs. Deviation` = sprintf("%.3f (%.3f)", median(mean_abs_error), IQR(mean_abs_error)),
        `Log Abs. Deviation` = sprintf("%.2f (%.2f)", median(mean_log_abs_error), IQR(mean_log_abs_error))
    ) |>
    # Apply kable formatting (optional, depends on your workflow)
    knitr::kable(escape = FALSE, booktabs = TRUE, align = c("l"))

@@

::: {#tbl-s2-agg apa-note="Summary statistics by reference class condition, based on participant-level averages. Exact Match Rate and Close Match Rate represent the mean percentage of trials (per participant) where the proposed reduction exactly matched or closely matched (±2% tolerance) the target goal (standard deviations in parentheses), respectively. Abs. Deviation and Log Abs. Deviation reflect the median (IQR) of each participant’s absolute deviation from the target goal."}
